### 2020-01-02 sections/introduction.tex
continues work on fleshing out paper. Adds discussion. Almost at 1st draft.

% BEGIN HERE
Spiral structure is present in a majority of massive galaxies (\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}), yet the formation mechanisms through which spiral structure originates are still hotly debated. Spirals themselves are as diverse as the theories proposed to govern their evolution; ranging from the quintessential pair of well-defined arcs of the grand design spiral, to the patchy and fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral. These variatons on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}). Our current understanding of the mechanisms which drive spiral growth and evolution suggest that each of the different forms of spiral galaxy may be triggered primarily by different processes. Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as suggested by Manifold theory, \citealt{2011hsa6.conf..548R}), or be obeying quasi-stationary density wave theory, in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}. Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and reccurent in nature \citep{1966ApJ...146..810J}. However it is recognised that no two methods of spiral formation are mutually exclusive.

@@

Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (\citealt{2015A&A...582A..86H}), fourier analysis (i.e. \textsc{2DFFT}, \citealt{2012ApJS..199...33D}), texture analysis (i.e. SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers \citep{2017MNRAS.472.2263H}. One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} (\comment{Lingard et al, in prep}). These methods attempt to localize light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization. This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}. \comment{Lingard et al, in prep} successfully solve this problem through the use of citizen science to provide the starting point for a computational fit.

A common assumption when measuring galaxy pitch angle is that observed spiral arms have a constant pitch angle, these spirals are known as logarithmic spirals and are described by

\begin{equation}
r = A\,e^{\theta\tan\phi}.
\end{equation}

One method used to obtain a pitch angle of a galaxy from identified arm segments is by taking the weighted mean of their pitch angles (as used by \textsc{SpArcFiRe}, \citealt{2014ApJ...790...87D}), where weighting is determined by the length of the arc segment, with longer arcs having higher weights, i.e. for a galaxy where we have identified $N$ arm segments, each with length $L_i$ and pitch angle $\phi_i$

@@

One notable drawback of the use of length-weighted arms is the sensitivity of the end result to the number and quality of the spiral arm segments identified \comment{...}

This paper makes use of the photometric models obtained through the \textit{Galaxy Builder} citizen science project for the 196 spiral galaxies present in \comment{Lingard et al, in prep}. In this paper we focus on the use of measured spiral tightness (quantified using pitch angle, \citealt{1987gady.book.....B}) as a probe into the dynamical mechanisms governing a spiral galaxy's evolution. We make use of Bayesian hierarchical modelling to measure galaxy pitch angle from the spiral arm clusters producted by \textit{Galaxy Builder}. We test for signs of spiral arm winding using the predictions derived by \citealt{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$) and conclude using a marginalized Anderson-Darling test that we \comment{cannot unilaterally reject winding of this form at the 1\% level}. Section \ref{section:morphology_comparision} examines the correlation between pitch angle and bulge size, implied by the Hubble sequence, and bar strength, implied by Manifold theory, and find \comment{no significant correlation}.

>>>>

% BEGIN HERE
Spiral structure is present in a majority of massive galaxies (\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}) yet the formation mechanisms through which spiral structure originates are still hotly debated. Spirals are as diverse as the theories proposed to govern their evolution; from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral. These variatons on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}). Our current understanding of the mechanisms which drive spiral growth and evolution suggest that each of the different forms of spiral galaxy may be triggered primarily by different processes. Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as suggested by Manifold theory, \citealt{2011hsa6.conf..548R}), or be obeying quasi-stationary density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}. Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and reccurent in nature \citep{1966ApJ...146..810J}. It is recognised that methods of spiral formation are not mutually exclusive. There is also debate as to the evolution of spiral structure: whether spirals maintain their shape over many rotational periods (assumed by QSDW theory, \citealt{1964ApJ...140..646L}), or ``wind up'' with the differential rotation of the disc (as found by \citealt{2019MNRAS.487.1808M}).

@@

Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (\citealt{2015A&A...582A..86H}), fourier analysis (i.e. \textsc{2DFFT}, \citealt{2012ApJS..199...33D}), texture analysis (i.e. SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers \citep{2017MNRAS.472.2263H}. One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} (\comment{Lingard et al, in prep}). These methods attempt to localize light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization. This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}. \comment{Lingard et al, in prep} propose a solution to this problem through the use of citizen science to provide the starting points for computational fitting.

A common assumption when measuring galaxy pitch angle is that observed spiral arms have a constant pitch angle. Spirals of this kind are known as logarithmic spirals and are described by

\begin{equation}
r = A\,e^{\theta\tan\phi},
\end{equation}

where $\phi$ is the arm's pitch angle. One method used to obtain a pitch angle of a galaxy is to fit logarithmic spirals to individually identified arm segments and take the weighted mean of their pitch angles (which often vary by upwards of $10^\circ$, \citealt{2014ApJ...790...87D}). Weighting is determined by the length of the arc segment, with longer being assigned higher weights, i.e. for a galaxy where we have identified $N$ arm segments, each with length $L_i$ and pitch angle $\phi_i$

@@

A notable drawback of length-weighted pitch angle is sensitivity to the number and quality of the spiral arm segments identified; \citet{2017MNRAS.472.2263H} found that only 15\% of the arm segments identified by a leading algorithm (\textsc{SpArcFiRe}) were identified as ``good'' matches to real spiral arms by citizen science classifiers.

Fourier analysis in one-dimension (as performed by \comment{...}) and two-dimensions (\comment{...}) is another widely used method of computationally obtaining galaxy pitch angles. Two-dimensional fourier methods generally decompose a deprojected image of a galaxy into a superpositions of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D}, and reports the pitch angle with the highest amplitude as the galaxy's pitch angle. It is uncertain how the large observed variation between pitch angles of individual arms impacts this measurement.

This paper makes use of the photometric models obtained through the \textit{Galaxy Builder} citizen science project for the 196 spiral galaxies present in \comment{Lingard et al, in prep}. In this paper we focus on the use of measured spiral tightness (quantified using pitch angle, \citealt{1987gady.book.....B}) as a probe into the dynamical mechanisms governing a spiral galaxy's evolution. We make use of Bayesian hierarchical modelling to measure galaxy pitch angle from the spiral arm clusters producted by \textit{Galaxy Builder}.

Section \ref{section:spiral_winding} investigates spiral arm winding using the test derived by \citealt{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$) and concludes using a marginalized Anderson-Darling test that we \comment{cannot unilaterally reject winding of this form at the 1\% level}. Section \ref{section:morphology_comparision} examines the correlation between pitch angle and bulge size implied by the Hubble sequence and reported by \comment{citation}, and pitch angle and bar strength implied by Manifold theory, and find \comment{no significant correlation}.

### 2020-01-02 sections/method.tex
continues work on fleshing out paper. Adds discussion. Almost at 1st draft.

% !TEX root = ../main.tex
\subsection{The Galaxy Sample}
The galaxies selected analysed in this paper are the 198 galaxies from \comment{Lingard et al, in prep.}. We combine classifications of galaxies which were repeated in the \textit{validation subset} with the original classifications, and perform clustering and point cleaning as detailed in \comment{Lingard et al, in prep.}, and remove any galaxies for which no arms were identified. This results in a hierarchiacal data structure of 109 galaxies and 250 spiral arms, with 307,861 points in polar coordinates, which are scaled such that the radius of each spiral arm has unit maximum.

@@

Assume we can model spiral arms as a logarithmic spiral, described by

@@

Assume the pitch angle of a galaxy's spiral arms are drawn from a Normal distribution, truncated between 0 and 90, centred on some value we will call the galaxy's pitch angle, $\phi_\mathrm{gal}$. This dispersion in arm pitch angle has some measure of spread, $\sigma_\mathrm{gal}$, which we will assume is the same in galaxies:

@@

We choose priors on $\phi_\mathrm{gal}$ and $\sigma_\mathrm{gal}$ of

@@

\begin{figure}
  \includegraphics[width=8cm]{plots/plots_n109d1000t500/model.pdf}
  \caption{The model used for galaxy pitch angle measurement for the sample.}
  \label{fig:ad-cot-test}
\end{figure}

To perform inference, we make use of the No-U-Turn-Sampler (NUTS), implemented in PYMC3\footnote{\url{https://docs.pymc.io/}}, an open source probabilistic programming framework written in python. \comment{does PYMC3 have a citation?}

>>>>

% !TEX root = ../main.tex
\subsection{The Galaxy Sample}
The galaxies selected analysed in this paper are the 198 galaxies from \comment{Lingard et al, in prep.}. We combine classifications of galaxies which were repeated in the \textit{validation subset} with the original classifications, and perform clustering and point cleaning as detailed in \comment{Lingard et al, in prep.}. We remove any galaxies for which no arms were identified. This results in a hierarchiacal data structure of 109 galaxies, 250 spiral arms and 307,861 points in polar coordinates. We apply a scaling to the radius to unit variance (which does not impact log spiral pitch angle due to its scale invariance).

@@

We model a spiral arm as a logarithmic spiral, described by

@@

Assume the pitch angle of a galaxy's spiral arms are drawn from a Normal distribution, truncated between 0 and 90 degrees, centred on some value we will call the galaxy's pitch angle, $\phi_\mathrm{gal}$. This dispersion in arm pitch angle has some measure of spread, $\sigma_\mathrm{gal}$, which we will assume is the same in galaxies:

@@

We choose hyperpriors on $\phi_\mathrm{gal}$ and $\sigma_\mathrm{gal}$ of

@@

\begin{figure}
  \includegraphics[width=8cm]{plots/plots_n109d1000t500/model.pdf}
  \caption{The model used for galaxy pitch angle measurement for the sample.}
  \label{fig:pymc3-model}
\end{figure}

To perform inference, we make use of the No-U-Turn-Sampler (NUTS, \citealt{2011arXiv1111.4246H}), implemented in PYMC3\footnote{\url{https://docs.pymc.io/}}, an open source probabilistic programming framework written in python \citep{pymc3_paper}.

### 2020-01-02 sections/results.tex
continues work on fleshing out paper. Adds discussion. Almost at 1st draft.

\subsection{Spiral Winding}

In order to test the possible progenitor distribution of our estimated galaxy pitch angles, we repeatedly perform an Anderson-Darling test over each draw present in the MCMC trace, resulting in a distribution of Anderson-Darling statistics. We will refer to this test as the \textit{marginalized Anderson-Darling test}.

We perform the marginalized Anderson-Darling test on a distribution uniform in $\cot\phi$ between the limits present in \citet{2019arXiv190910291P} ($1.00 < \cot\phi < 4.75$, or roughly $11.9 < \phi < 45.0$). The resulting distribution of Anderson-Darling statistics can be seen in Figure \ref{fig:ad-cot-test}. We observe that we reject the null hypothesis at the 1\% level for only 92\% of the possible realizations of galaxy pitch angle. Therefore we cannot unilaterally reject winding of the kind described by \citet{2019arXiv190910291P}, though .

\begin{figure*}
  \includegraphics[width=17.7cm]{plots/cot_uniform_marginalized_tests.pdf}
  \caption{The results of a marginalized Anderson-Darling test (top panel), with confidence intervals shown, and marginalized Kolmogorov-Smirnov test p-values (lower panel).}
  \label{fig:ad-cot-test}
\end{figure*}

\subsection{Dependence of pitch angle on Galaxy Morphology}
\label{section:morphology_comparision}
\subsubsection{Pitch angle vs. Bulge size}
We see no correlation between galaxy pitch angle derived from the \textit{hierarchial normal model} and GZ2's \textit{pbulge}, which is widely viewed as a good measure of bulge size.

\subsubsection{Pitch angle vs. Bar Strength}
We see no correlation between galaxy pitch angle derived from the \textit{hierarchial normal model} and GZ2's \textit{pbar}, which is widely viewed as a good measure of bar strength, and therefore a measure of the torque applied on the disc gas.

A marginalized Anderson-Darling test does not find that the samples were drawn from different distributions (statistic of -0.91).

>>>>

\subsection{Spiral Winding}
\label{section:spiral_winding}
In order to test the possible progenitor distribution of our estimated galaxy pitch angles, we repeatedly perform an Anderson-Darling test over each draw present in the MCMC trace, resulting in a distribution of Anderson-Darling statistics. We will refer to this test as the \textit{marginalized Anderson-Darling test}. We make use of the Kolmogorov-Smirnov test in a similar manner for comparison.

We perform the marginalized Anderson-Darling test for a potential source distribution uniform in $\cot\phi$ between the limits present in \citet{2019arXiv190910291P} ($1.00 < \cot\phi < 4.75$, or roughly $11.9^\circ < \phi < 45.0^\circ$). The resulting distribution of Anderson-Darling statistics can be seen in Figure \ref{fig:ad-cot-test}. We observe that we reject the null hypothesis at the 1\% level for only 86\% of the possible realizations of galaxy pitch angle, therefore with our sample and methodology we cannot unilaterally reject winding of the kind described by \citet{2019arXiv190910291P}.

\begin{figure*}
  \includegraphics[width=17.7cm]{plots/cot_uniform_marginalized_tests.pdf}
  \caption{The results of a marginalized Anderson-Darling test (top panel), with confidence intervals shown, and marginalized Kolmogorov-Smirnov test p-values (lower panel). Moving rightwards on the x-axis implies greater confidence in rejecting the null hypothesis.}
  \label{fig:ad-cot-test}
\end{figure*}

\subsection{Dependence of pitch angle on Galaxy Morphology}
\label{section:morphology_comparision}
\subsubsection{Pitch angle vs. Bulge size}
We see no correlation between galaxy pitch angle derived from the \textit{hierarchial normal model} and Galaxy Zoo 2's debiased \citep{Willett2013:1308.3496v2} \textit{pbulge}, which has been shown to be a good measure of bulge size.

We separate our sample into ``disc-dominated galaxies'' and ``obvious bulge galaxies'' using the debiased fractions from Galaxy Zoo 2 following \citet{2017MNRAS.469.3363K}, defining \textit{no bulge + just noticeable $>$ obvious + dominant} for the former and the converse for the latter. A marginalized two-sample Anderson-Darling test \citep{doi:10.1080/01621459.1987.10478517} does not find strong evidence that the samples were drawn from different distributions; we reject the null hypothesis at the 1\% level for only 8\% of the samples.

\subsubsection{Pitch angle vs. Bar Strength}
We see no correlation between galaxy pitch angle derived from the \textit{hierarchial normal model} and Galaxy Zoo 2's debiased \textit{pbar}, which is widely viewed as a good measure of bar strength, and therefore a measure of the torque applied on the disc gas.

Separating the sample based off of $\mathrm{\textit{pbar}} > 0.5$, and restricting to galaxies with more than 10 classifications for \textit{pbar} (as performed by \citealt{2011MNRAS.411.2026M} and \citealt{2017MNRAS.469.3363K}) and performing a marginalized two-sample Anderson-Darling test does not find that the samples were drawn from different distributions.

\begin{figure*}
  \includegraphics[width=17.7cm]{plots/bulge_bar_test_results.pdf}
  \caption{The results of marginalized two-sample Anderson-Darling tests examining whether pitch angles for Bulge-dominated and Disc-dominated galaxies are drawn from the same distribution (top panel), and the results of the same test for strongly-barred vs unbarred galaxies (bottom panel).}
  \label{fig:ad-morphology-test}
\end{figure*}
